InferentialStats

### 2018-08-29 E2/analysis/dev/Chapter1.Rmd
right

---
title: "validateE2data"
author: "Alex Holcombe"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

@@

```{r}
readInAllFiles<- function(rawDataPath) {

  files <- dir(path=rawDataPath,pattern='.mat')  #find all data files in this directory
  dfAll<-data.frame()
  for (i in 1:length(files)) { #read in each file
    fileThis<- file.path(rawDataPath,files[i])
    rawDataLoad=tryCatch(
      readMat(fileThis),
      error=function(e) {
        stop( paste0("ERROR reading the file ",fileThis," :",e) )
      } )
    apparentSubjectName <- strsplit(files[i],split="_")[[1]][1]
    subjectName<- rawDataLoad$participantID[[1]]
    if (apparentSubjectName != subjectName) {
      stop( paste0("WARNING apparentSubjectName",apparentSubjectName," from filename does not match subjectName in data structure",subjectName) )
    }
    else { cat(paste0(' Subject ',apparentSubjectName,' read in')) }
    rawDataLoad$file <- files[i]

    dfThis<- turnMATintoMeltedDataframe(rawDataLoad)

    tryCatch(
      dfAll<-rbind(dfAll,dfThis), #if fail to bind new with old,
      error=function(e) { #Give feedback about how the error happened
        cat(paste0("Tried to merge but error:",e) )
      } )
  }
  return(dfAll)
}
```

```{r readRawData}

E<- readInAllFiles(rawDataPath)

#Calculate the serial position error
E$SPE<- E$respSP - E$targetSP

```

Why are there NaNs for targetSP? Because that's when only one target (targetSide==1 or ==2)

```{r}
table(E$subject, E$targetSP,useNA="always")
```

Show how targetSP works
```{r}
library(dplyr)
En <- E
En$letterSeq<- NULL #otherwise can't use dplyr
CU <- En %>% filter(subject=="CU")
table(CU$target,CU$targetSide, useNA="always")
#targetSide ==0  means targets on both sides 
table(CU$targetSide, CU$targetSP, useNA="always")
```

The below works now that turnMATintoMeltedDataframe takes care of the list elements

```{r}
#CUall <- E[ ,E$subject=="CU" ]
CUall <- E %>% filter(subject=="CU")
i<- E$subject == "CU"
E[i,] #previously didn't work. You can't slice a dataframe that has a list (letterSeq) as a member.
subset(E, subject=="CU")
j<- En$subject == "CU"
EnCU<- En[j,]
```

Load in one file and inspect it
```{r}

source(file.path("turnMATintoMeltedDataframe.R"))

fileThis<- file.path(rawDataPath, "CU_18-05-10_1.mat")

rawDat=tryCatch(
  readMat(fileThis),
  error=function(e) {
    stop( paste0("ERROR reading the file ",fileThis," :",e) )
  } )
str(rawDat)
CUmelted<- turnMATintoMeltedDataframe(rawDat)

```
Somehow, what would seem to be the second trial has NaN for lots of stuff

 $ allLetterOrder         : num [1:216, 1:2, 1:16] 13 6 9 1 10 16 11 1 12 8 ...
 $ allLetterOrientations  : num [1:216, 1:16] 1 0 1 0 0 0 1 0 1 0 ...
 $ allTargets             : num [1:216, 1:2] 5 NaN 6 8 7 7 6 6 6 8 ...
 $ allTargetOrientations  : num [1:216, 1:2] 0 NaN 1 1 0 1 1 1 0 0 ...
 $ allResponses           : num [1:216, 1:2] 9 NaN 1 3 5 11 5 3 8 11 ...
 $ allResponseOrientations: num [1:216, 1:2] 0 NaN 0 0 0 1 0 0 0 0 ...
 $ allResponseOrder       : num [1:216, 1] 2 2 1 2 2 1 2 1 1 1 ...
 $ allRTs                 : num [1:216, 1:2] 12.78 NaN 1.81 6.19 10.19 ...

Inspect another one 
```{r}
fileThis<- file.path(rawDataPath, "BC_18-01-25_1.mat")

rawDat=tryCatch(
  readMat(fileThis),
  error=function(e) {
    stop( paste0("ERROR reading the file ",fileThis," :",e) )
  } )
str(rawDat)

```

Sanity visual check that see histogram.
```{r sanity visual check}

sanityVisualCheck<-TRUE
if (sanityVisualCheck) {
  library(ggplot2)
  g=ggplot(E, aes(x=SPE))
  #plot data
  g<-g+geom_histogram(binwidth=1)
  show(g)

  h<-ggplot(E,aes(x=SPE)) + geom_histogram(binwidth=1)
  show(h)
}

```

Meaning of condition variable
* 1 - upright
* 2 - inverted

Make condition names more sensible - stream (Left/Right) and orientation (Canonical/Inverted)
```{r}
E2<- E
names(E2)[names(E2) == 'target'] <-'stream'
names(E2)[names(E2) == 'condition'] <-'orientation'
E2$stream[ E2$stream==1 ] <- "Left"
E2$stream[ E2$stream==2 ] <- "Right"
#mutate condition to Orientation
E2$orientation[ E2$orientation==1 ] <- "Canonical"
E2$orientation[ E2$orientation==2 ] <- "Inverted"

```

Sanity-check some data by graphing.

Investigate whether pathological identicalness of inverted/canonical happens, that happened in E1 . Seems it doesn't

```{r}
dg<- E2 %>% filter(subject < "AC")

library(ggplot2)
g=ggplot(dg, aes(x=SPE)) + facet_grid(subject+orientation~stream) #,  scales="free_y")
minSPE<- -17; maxSPE<- 17  #But it's 16 items in stream, so shouldn't it be -15, +15?
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
show(g)
```

Save data
```{r}
#save dataframe into proper library data directory
backwards2_E2<-E2
saveRDS(backwards2_E2, file = "../Data/backwards2E2_rawDataFromMAT.rda", compress="bzip2")
```

Look at 1-target performance, understand that variable better because didn't exist in previous experiment I think.

The first table printed below could easily fool you into thinking that stream (Left, Right) and targetSide (1,2) are not consistent. We think that's because it records something for both sides (stream variable) even when there is only 1 target (targetSide == 1 or 2 rather than 0, which means two targets).

I Disaggregated allLetterOrientations - done by importE2data.Rmd which renames it targetStreamOrientations

Also maybe give targetSide values that make more common sense, like "dual", "left", and "right"

But the second table below shows that they make sense, as they are only inconsistent when allResponseOrientations==NaN.  But what the hell are those trials where  allResponseOrientations==NaN?
```{r}
library(dplyr)
E2$letterSeq <- NULL
one <- E2 %>% filter(numTargets==1)

table(one$stream,one$targetSide,useNA="always")

table(E2$stream,E2$targetSide,E2$numTargets, useNA="always")

table(one$stream,one$targetSide,one$orientation,useNA="always")

#table(one$stream,one$targetSide,one$allResponseOrientations,useNA="always")
```

>>>>

---
title: "Chapter1"
author: "Alex Holcombe"
date: "8/28/2018"
output: html_document
---

@@

# Chapter 1

This is chapter 1.

```{r}
1
```

### 2018-08-29 E2/analysis/inferentialStats.Rmd
right

@@

estimates<- read.csv(file="Results/backwards2E1_paramEstimates.csv", header=TRUE, sep=",")
estimates$X<- NULL #Not sure where X comes from but it's just a row number
```

Exclude Ss who failed likelihood ratio test in all four cases.
Create new column nPassed for in how many conditions they passed (out of 4).

```{r}
library(dplyr)

estimates<- estimates %>% mutate(mixModelFitsBetter = pLRtest<.05)

condtnVariableNames <- c("subject","stream", "orientation")

e<- estimates %>%  group_by(subject) %>%
  summarise(nPassed = sum(mixModelFitsBetter)) 
estimates<- merge(e,estimates)

estimates<- estimates %>% mutate( exclude = (nPassed==0)  )
```

Assess the size of the effect to inform Bayesian test of next experiment (probably use raw units)

```{r, echo=FALSE, message=FALSE}
require(ggplot2)
gg<-ggplot(estimates, aes(x=stream, y=efficacy)) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
show(gg)

library(data.table)

#The below works but drops many variables, so think about merging later
eff<- data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="efficacy")
eff$dv<-"efficacy"
lat<-data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="latency") 
lat$dv<-"latency"
pre<-data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="precision") 
pre$dv<-"precision"
all<-rbind(eff,lat,pre)
#I tried to do the above with a single command but makes an error even though StackOverflow example works
#https://stackoverflow.com/questions/37332076/from-long-to-wide-data-with-multiple-columns/37332345
# data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var=c("efficacy","latency"))
head(eff)
```

Calculate bias
```{r , fig.height=20, fig.width=5}
all$leftBias<- all$Left - all$Right
hh<-ggplot(all, aes(x=orientation, y=leftBias)) + facet_grid(.~dv) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
hh<-hh+stat_summary(fun.data="mean_cl_boot",geom="errorbar",width=.3,conf.int=.95,
                  width=5,size=1) 
show(hh)
```

>>>>

```{r}
dataPath <- file.path( "..","Data")
#Created by exclusions.Rmd
data<- readRDS( file.path(dataPath, "backwards2E2_dataWithEstimatesAndExcludeColumn.rda") ) 
data<- as_tibble(data)
```

Why is there no targetSP==NA in data? That's the rows reporting stream on opposite side. Eventually, use those to evaluate swaps. I don't know what happened to it, because it's created in 
exclusions.Rmd, which on line 67 creates the thing merged with estimates that is saved. 
filter( !is.na(targetSP) )

For inferential stats, don't care about individual trials, so extract parameter estimates from full tibble.

@@

estimates<- data %>% select(-targetStreamOrientations,-letterSeq) %>% #can't include list columns
  group_by(subject,targetSide,orientation,stream) %>% #in other words, average over trials 
  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })

library(readr)
resultsPath = "Results/"
write_tsv(estimates, file.path(resultsPath,"backwards2E2_paramEstimatesWithExcludeColumn.tsv"))

Plot data

```{r, echo=FALSE, message=FALSE}
require(ggplot2)
gg<-ggplot(estimates %>% filter(targetSide=="Both"), 
           aes(x=stream, y=efficacy, color=orientation)) + geom_point() + stat_summary(fun.y="mean",geom="point",size=8, alpha=.5) + theme_apa() +ggtitle('nobody excluded')
show(gg)

hh<-ggplot(estimates %>% filter(targetSide=="Both", exclude==FALSE), aes(x=stream, y=efficacy, color=orientation)) + 
  geom_point(position = position_dodge(width = 0.3)) + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5, position = position_dodge(width = 0.3)) + 
  theme_apa() +ggtitle('after exclusions') + 
  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = 0.3))

show(hh)

Graph the latency and sigma as well
BG, BJ, and AP have sigma near 4, clear outliers.

```{r, echo=FALSE, message=FALSE}
require(ggplot2)

hh<-ggplot(estimates %>% filter(targetSide=="Both", exclude==FALSE), aes(x=stream, y=latency, color=orientation)) + 
  geom_point(position = position_dodge(width = 0.3)) +  stat_summary(fun.y="mean",geom="point",size=4, alpha=.5, position = position_dodge(width = 0.3)) + 
  theme_apa() +ggtitle('after exclusions') + 
  stat_summary(fun.data =mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = 0.3))
show(hh)

ii<-ggplot(estimates %>% filter(targetSide=="Both", exclude==FALSE), aes(x=stream, y=precision, color=orientation)) + 
  geom_point(position = position_dodge(width = 0.3)) +  stat_summary(fun.y="mean",geom="point",size=4, alpha=.5, position = position_dodge(width = 0.3)) + 
  theme_apa() +ggtitle('after exclusions') + 
  stat_summary(fun.data =mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = 0.3))
show(ii)

Convert from long to wide so that both left and right stream estimates are in the same row, while going from wide to long for efficacy,latency,precision so can use facet_grid with them.

```{r, echo=FALSE, message=FALSE}

library(data.table)
#The below works but drops many variables, so think about merging later
#I guess what comes after the "~" is what goes from long to wide
eff<- data.table::dcast(estimates, orientation+subject+exclude+targetSide~stream,
                        value.var="efficacy")
eff$dv<-"efficacy"
lat<-data.table::dcast(estimates, orientation+subject+exclude+targetSide~stream,
                       value.var="latency") 
lat$dv<-"latency"
pre<-data.table::dcast(estimates, orientation+subject+exclude+targetSide~stream,
                       value.var="precision") 
pre$dv<-"precision"
all<-rbind(eff,lat,pre)
#I tried to do the above with a single command but makes an error even though StackOverflow example works
#https://stackoverflow.com/questions/37332076/from-long-to-wide-data-with-multiple-columns/37332345
# data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var=c("efficacy","latency"))
head(all)
```

Calculate and plot bias. 
```{r , fig.height=20, fig.width=5}
all$leftBias<- all$Left - all$Right
hh<-ggplot(all %>% filter(targetSide=="Both"), 
           aes(x=orientation, y=leftBias)) + facet_grid(dv~., scales="free") +
    stat_summary(fun.y="mean",geom="point",size=4, alpha=.5) + geom_point() +
  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.2) + theme_apa() +

show(hh)
```
